Vagueness and imprecise imitation in signalling games

### 2016-05-26 paper/version_final/paper.tex
Ran aspell spell-checker.

\title{Vagueness and imprecise imitation in signaling games}
% \author{Michael Franke \and Jos\'e Pedro Correia}
\date{}

@@

\begin{abstract}
  Signaling games are popular models for studying the evolution of meaning, but typical
  approaches do not incorporate vagueness as a feature of successful signaling.  Complementing
  recent like-minded models, we describe an aggregate population-level dynamic that describes a
  process of imitation of successful behavior under imprecise perception and realization of
  similar stimuli. Applying this new dynamic to a generalization of Lewis' signaling games, we
  show that stochastic imprecision leads to vague, yet by-and-large efficient signal use, and,
  moreover, that it unifies evolutionary outcomes and helps avoid suboptimal
  categorization. The upshot of this is that we see \emph{as-if}-generalization at an aggregate
  level, without agents actually generalizing.
\end{abstract}

@@

But there are other intriguing aspects about vagueness. One perplexing issue is how vagueness
could arise and be maintained in the first place. This is an apparent puzzle for functionalist
accounts that maintain that concepts and linguistic meanings evolved driven towards
efficiency. \citet{Lipman2009:Why-is-Language} argues that, in common interest signaling
situations, the existence of unclear borderline cases entails inefficiency of categorization or
communication, or at least no advantage. The challenge is then to explain how vagueness
can persist both: (i) under evolutionary pressure to be optimally discriminative, and (ii)
without undermining the possibility of evolving, learning and communicating with a meaningful
language. A number of authors have consequently tried to explain why vagueness evolved as
something that is itself useful
\citep[e.g.][]{Jaegherde-Jaegher2003:A-Game-Theoreti,Deemter2009:Utility-and-Lan,BlumeBoard2013:Intentional-Vag}.
Others have argued that vagueness is a natural byproduct of limitations in information
processing \citep[e.g.][]{FrankeJager2010:Vagueness-Signa} or of generalization in low-level
learning strategies \citep[e.g.][]{OConnor2013:The-Evolution-o}. This paper contributes to the
latter line of thought.

We believe that vagueness in language and thought may have many reasons, not just one. We focus
here on one \emph{a priori} plausible reason for why vagueness is natural and pervasive. The
idea is that vagueness is, at least in part, due to imprecision in the perception of similar
stimuli and imprecision in the realization of similar responses. On this view, vagueness in
language may be seen as a necessary sub-optimality due to limitations in another domain of
cognition, first and foremost perception. It could then be speculated that the whole system,
perception and language together, may be an almost optimal adaptation in a larger frame
of reference, e.g., if we take into account the metabolic costs for increased perceptual
accuracy. We will not engage in such speculation here. Rather we will explore the consequences
of perceptual limitations on processes of meaning evolution in a suitable formal framework. In
other words, in order to address Lipman's challenge seriously, and not just hand-wave it away
by appeal to the naturalness of vagueness, it needs to be spelled out how exactly confusability
of stimuli could come into play in a process of meaning evolution and how it could lead to
vague but by-and-large informative signal meaning. The formal model that this paper introduces
does exactly that. But it also does more. We find that confusability of stimuli can regularize
and systematize evolving meaning. This suggests a possibly advantageous side-effect that a
natural cause of vagueness may have and that might compensate some of the disadvantages that
vagueness may have for communicative efficiency \citep[c.f.][]{OConnor2013:The-Evolution-o}.

@@

The next section introduces the background against which the work presented here can be
appreciated. Section~\ref{sec:repl-diff-dynam} introduces a generalization of the replicator
dynamic which is derived from the idea that agents imitate other agents' behavior while
possibly confusing similar states. This imprecise imitation dynamic is explored in
Section~\ref{sec:exploring-rdd}. Section~\ref{sec:discussion} reflects and compares our
approach to others. Appendix~\ref{sec:formal-stuff} provides formal detail.

@@

Signaling games, as introduced by \citet{Lewis_1969:Convention}, have
a sender and a receiver. The sender knows the true state of the world,
but the receiver does not. The sender can select a signal, or message,
to reveal to the receiver, who then chooses an act. In Lewis' games,
if the receiver chooses the act that corresponds to the actual state,
the play is a success, otherwise a failure. Certain regular
combinations of sender signaling and receiver reaction make messages
meaningful, in the sense that their use is correlated systematically
to certain states or acts. To investigate the conditions under which
such meaning-generating behavior can evolve is a highly interesting
topic that we are only beginning to fully understand
\citep[e.g.][]{Warneryd1993:Cheap-Talk-Coor,BlumeKim1993:Evolutionary-St,Huttegger2007:Evolution-and-t,Pawlowitsch2008:Why-Evolution-d,Barrett2009:The-Evolution-o,HutteggerSkyrms2010:Evolutionary-Dy,Skyrms2010:Signals}.

Similarity-maximizing (short: sim-max) games are a variation of Lewis' games where the
receiver's actions are equated with the state space (one can think of the actions as choosing
states) and different states are allowed to be more or less similar to one another. While
Lewis' games treated communicative success as a matter of black and white, sim-max games allow
for shades of gray: the more similar the receiver's interpretation is to the actual state, the
better. Signaling games with utility-relevant similarities in the state space are fairly
standard in economics
\citep[e.g.][]{Spence1973:Job-market-sign,CrawfordSobel1982:Strategic-Infor}, but have received
particular attention in a more philosophical context for reasons that will become clear
presently.

Formally, a sim-max game consists of a set of states $\States$, a set of messages $\Messgs$
typically with much fewer messages than states, a probability distribution
$\Pr \in \Delta(\States)$ such that $\Pr(\state)$ gives the probability that state $\state$
occurs, a similarity metric\footnote{A metric is a function of distance between any two points
  in a given space. It should satisfy certain axioms that ensure behavior that one would
  intuitively expect from the words ``similarity'' and ``distance'' alone, but these details do
  not matter for the purposes of this paper. The assumption that state similarity forms a
  metric is conceptually loaded, but we follow the literature here and conceive of it as a
  first and pragmatic simplification, possibly to be dispensed with later.}
  % : similarity should always be non-negative, zero only for the distance
  % between an element and itself, symmetric, and satisfy the triangle inequality that
  % $\similarity(x,z) \le \similarity(x,y) + \similarity(y,z)$
on states $\similarity \mycolon \States \times \States \rightarrow \mathds{R}$ such that
$\similarity(\mystate{1}, \mystate{2})$ is the (physical) similarity between $\mystate{1}$ and
$\mystate{2}$, and a utility function
$\Utils \mycolon \States \times \States \rightarrow \mathds{R}$ such that
$\Utils(\mystate{1}, \mystate{2})$ is the payoff for sender and receiver for a play with actual
state $\mystate{1}$ and receiver interpretation $\mystate{2}$. We identify the receiver's acts
with the states of the world, so that the game is one of guessing the actual state, so to
speak. For the modeling purposes of this paper, we make the simplifying assumption that
$\States$ is a set of points in Euclidean space, whose closeness to each other tracks physical
similarity. Perceived similarity, where it is necessary, would be a monotonic function of
physical similarity. Likewise, the utility function should be a monotonically increasing
function of physical similarity.\footnote{Section~\ref{sec:setting-stage} motivates particular
  choices of similarity and utility functions that we will explore in more detail.}
Non-probabilistic sender behavior can be represented by a pure strategy
$\Spure \in \Messgs^\States$, which deterministically defines which message would be used for
each state. Similarly, a pure receiver strategy is a function $\Rpure \in \States^\Messgs$.

@@

It is intuitive to think that linguistic and conceptual categories are orderly in such a
manner.  For example, if you consider two people `tall', one with a height of 2m and another with a
height of 2.2m, it would be difficult to defend not considering a person with a height
of 2.1m `tall' as well. Another example where this intuition is additionally supported by empirical data is that of color categorization. The World Color Survey
project~\citep{CookKay2005:The-World-Color,WCS} collected color naming data for 110 unwritten
languages of 45 language families.  For the great majority of these languages a pattern can be
observed: basic color terms are by and large convex~\citep{Regier07,Jager2010}. It
is premature to argue that these observations can be extended to \emph{all} cases of
categorization. However, for the cases where it does apply, the result of
\citet{JagerMetzger2011:Voronoi-Languag} is interesting because it demonstrates that signaling
can impose this kind of orderly categories on a metric space without that being the ulterior
purpose of it all.%
\footnote{For the concrete case of color categorization, see also
  \citet{JagerRooijvan-Rooij2007:Language-Struct} and \citet{Correia2015}}
Finally, these considerations are in line with a prominent
school of thought in comparative linguistics which also assumes that more abstract conceptual domains
(e.g., spatial-topological relations, temporal reference, or the meanings expressible by
indefinite pronouns) are preferably carved up by the languages of the world in such a way
that meanings are connected regions on a so-called \emph{semantic map}
\citep[e.g.][]{Croft2003:Typology-and-Un,Haspelmath2003:The-geometry-of,LevinsonMeira2003:Natural-concept}.

There are at least two potential ways of interpreting the signaling setup.
Sender and receiver can be distinct
entities, whose purpose is to communicate effectively about the actual
state. In that case, evolving Voronoi languages would explain why
\emph{linguistic categories} are well-behaved and orderly in the way
they appear to be. More abstractly, sender and receiver can also be
thought of as distinct modules in a single system, where the first
module must discretize the information it is fed by selecting a small
sample of, suggestively, category labels. These are passed to a second
module that tries to decode the original information. In this case,
evolving Voronoi languages would explain why \emph{conceptual
  categories} are well-behaved and orderly in the way that they appear
to be.
Seen in this light, sim-max games may provide a foundation to approaches in cognitive semantics
that rely on the notion of conceptual spaces.
\citet[][70--77]{Gardenfors2000:Conceptual-Spac}, for example, has prominently argued that
natural categories are convex regions in conceptual space. If the conceptual space has a
suitable metric, convex categories can be derived from a set of prototypes. The category
corresponding to prototype $p$ is the set of points that are more similar to $p$ than to any
other. In this way, \citeauthor{Gardenfors2000:Conceptual-Spac} argues, an efficient categorization
system can be obtained: storing the prototypes lets us recover the categories without having to
store each category's extension. However, what is left unexplained is where the prototypes
come from, and why we would not see just any distribution of prototypes as an equally efficient
classification system. This is where sim-max games can contribute a principled approach to
deriving, in an independent way, not only convex categories but also prototypical exemplars
belonging to them.  These ideas, and more, are developed further by
\citet{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag,OConnor2014-OCOEPC},
among others.

@@

This outline of an approach to categorization using sim-max games
leaves many problems unaddressed. One of them is that natural
categories for continuously variable stimuli usually do not have
unique, point-valued prototypes and clear category boundaries. We
would like to account for the possibility of such vagueness, in
particular: (i) clear positive examples of a vague category should
show a gradient transition to clear negative examples;
(ii) prototypes should likewise be gradient regions, peaking at the
center of the vague category they represent.

@@

Alternative approaches are taken by, for example, \citet{FrankeJager2010:Vagueness-Signa} and
\citet{OConnor2013:The-Evolution-o}, who show different ways how the above desiderata can be
met by evolving strategies in sim-max games. To illustrate what vague signaling would look
like, let us briefly consider a sim-max game with six equiprobable states and two messages and
what its equilibria would be like \citep[see][for a more thorough discussion of equilibria of
sim-max games]{Jager2007:The-Evolution-o,OConnor2013:The-Evolution-o}. We assume that utilities
are linearly decreasing with decreasing similarity. Figure~\ref{fig:EquilibriaVagueness} shows
three pairs of sender and receiver strategies. States are arranged according to their
similarity: the closer they are to each other, the more similar they are. The pair in
Figure~\ref{fig:ineffNonEq} is not an equilibrium, because the sender's non-convex use of
signals is suboptimal given the receiver's behavior. Namely, if \mymessg{1} is interpreted as
\mystate{2} and \mymessg{2} as \mystate{4}, then the sender would get a higher payoff from
sending \mymessg{1} in \mystate{3} than from sending \mymessg{2}, because (by assumption)
\mystate{3} is more similar to \mystate{2} than \mystate{5} is. In contrast,
Figure~\ref{fig:efficientEq} shows a maximally efficient equilibrium. This is a partial pooling
equilibrium in the sense that the speaker uses the same message for several states. Partial
pooling equilibria can be less inefficient than other non-pooling equilibria, if there are
enough messages. The strategy pair of Figure~\ref{fig:efficientEq} is maximally efficient for
the two-message case. So, while partial pooling may entail inefficiency in some sense, and
while partial pooling can hamper the evolution of maximally efficient signal use
\citep[e.g.][]{Hutteger:2007_Evol_Indicatives_Imperatives,Pawlowitsch2008:Why-Evolution-d,HutteggerSkyrms2010:Evolutionary-Dy},
this is orthogonal to our concerns about vagueness. A regular and natural yet vague signal use
would look like the pair in Figure~\ref{fig:vagueUse}. This is not an equilibrium, but it gets
close, so to speak. It shows smooth transitions across similar states at the boundaries of
categories and across acts around the most prototypical instances (as indicated by decreasing
thickness of arrows in Figure~\ref{fig:vagueUse}).

@@

Our conceptual starting point is the widely shared conviction that language is shaped by at
least two forces, which may, on occasion, pull in opposite direction. On the one hand, there is
functional pressure towards efficient communication. On the other hand, there is systematic
error, noise or imprecision in the transmission of linguistic behavior, knowledge or traits. As
an example for the latter, consider a child learning syntactic rules from a parent
generation. The child must infer these unobservable rules from observable speech. Inductive
biases may influence which syntactic rules are likely to be inferred from (finite) parental
input. Over the course of many generations, the effects of such biases can, in a manner of
speaking, ``accumulate'' and lead to surprising results, such as the evolution of compositional
form-meaning mappings or the use of regular recursive syntactic structure. This can happen when
the effects of tranmission biases are isolated, as in iterated learning models
\citep{KirbyHurford2002:The-Emergence-o,SmithKirby2003:Iterated-Learni,GriffithsKalish2007:Language-Evolut,KirbyGriffith2014:Iterated-Learni},
or when they interact with functional pressure towards efficient communication, e.g., as
formalized in the replicator mutator dynamic
\citep[e.g.][]{NowakPlotkin2000:The-Evolution-o,NowakKomarova2001:Evolution-of-Un}.

The emphasis of previous models that studied the effects of transmission infidelity on the
evolution of language has been on inductive biases and the systematicity, compression and
regularization that they can introduce. Here, we would like to show that transmission noise of
a different kind can lead to regularization as well and also give rise to vague meaning. We
find that shared perceptual biases that perturb the transmission of successful signaling
behavior can regularize, facilitate and accelerate the evolution of meaning conventions, albeit
at the cost of vagueness. Concretely, we formalize the expected change in the behavior of a
population of agents that try to imitate other agents' signaling behavior. We assume, however,
that both realization of behavior, as well as observability of others' behavior, is
systematically perturbed by noise. The resulting population-level dynamic generalizes the
replicator dynamic \citep{TaylorJonker1978:Evolutionary-St} in its interpretation as a cultural
evolutionary dynamic based on imitation
\citep[e.g.][]{Helbing1996:A-Stochastic-Be,Schlag1998:Why-Imitate-and}. The inclusion of
confusability of stimuli does not undermine the possibility of evolving communicative signaling
behavior. Instead, it leads to the evolution of vague meanings. It moreover accelerates the
emergence of communicative signaling because it unifies and regularizes evolutionary outcomes,
making it appear as if agents were applying inductive biases or generalizing over partial
observations, when in truth this is the sole effect of confusion of similar stimuli.

@@

Signaling agents can adapt their dispositions to act, given some feedback about their past
success, in multiple ways. Usually, we would assume that changes in behavior should, at least
on average, lean towards increasing chances of communicative success. Such behavioral
adaptations can be described at different levels of abstraction. At the level of individual
agents, we can picture a more or less idealized process of how each agent adapts dispositions
for future actions based on various pieces of information available to the agent. More
abstractly, at the level of a population of agents, we can describe how average behavioral
dispositions will evolve. The population-level perspective abstracts over small stochastic
fluctuations and zooms in on the general tendency or direction of evolution that ensues from
behavior at the agent-level.

To better understand the interaction of confusability of stimuli and selective pressure towards
successful communication, we look at a population-level dynamic that describes the most likely
evolutionary path of a population of signaling agents who imitate other agents' behavior but
are liable to confuse states for one another. In the special limiting case where state
confusability vanishes, the process is just the well-known replicator dynamic
\citep{TaylorJonker1978:Evolutionary-St}, which should therefore be briefly reviewed first.

\subsection{Replicator dynamic in behavioral strategies}
\label{sec:repl-dynam-behav}

>>>>

\title{Vagueness and imprecise imitation in signalling games}
% \author{Michael Franke \and Jos\'e Pedro Correia}
\date{}

@@

\begin{abstract}
  Signalling games are popular models for studying the evolution of meaning, but typical
  approaches do not incorporate vagueness as a feature of successful signalling.  Complementing
  recent like-minded models, we describe an aggregate population-level dynamic that describes a
  process of imitation of successful behaviour under imprecise perception and realisation of
  similar stimuli. Applying this new dynamic to a generalisation of Lewis' signalling games, we
  show that stochastic imprecision leads to vague, yet by-and-large efficient signal use, and,
  moreover, that it unifies evolutionary outcomes and helps avoid sub-optimal
  categorisation. The upshot of this is that we see \emph{as-if}-generalisation at an aggregate
  level, without agents actually generalising.
\end{abstract}

@@

But there are other intriguing aspects about vagueness. One perplexing issue is how vagueness
could arise and be maintained in the first place. This is an apparent puzzle for functionalist
accounts that maintain that concepts and linguistic meanings evolved driven towards
efficiency. \citet{Lipman2009:Why-is-Language} argues that, in common interest signalling
situations, the existence of unclear borderline cases entails inefficiency of categorisation or
communication, or at least no advantage. The challenge is then to explain how vagueness
can persist both: (i) under evolutionary pressure to be optimally discriminative, and (ii)
without undermining the possibility of evolving, learning and communicating with a meaningful
language. A number of authors have consequently tried to explain why vagueness evolved as
something that is itself useful
\citep[e.g.][]{Jaegherde-Jaegher2003:A-Game-Theoreti,Deemter2009:Utility-and-Lan,BlumeBoard2013:Intentional-Vag}.
Others have argued that vagueness is a natural byproduct of limitations in information
processing \citep[e.g.][]{FrankeJager2010:Vagueness-Signa} or of generalisation in low-level
learning strategies \citep[e.g.][]{OConnor2013:The-Evolution-o}. This paper contributes to the
latter line of thought.

We believe that vagueness in language and thought may have many reasons, not just one. We focus
here on one \emph{a priori} plausible reason for why vagueness is natural and pervasive. The
idea is that vagueness is, at least in part, due to imprecision in the perception of similar
stimuli and imprecision in the realisation of similar responses. On this view, vagueness in
language may be seen as a necessary sub-optimality due to limitations in another domain of
cognition, first and foremost perception. It could then be speculated that the whole system,
perception and language together, may be an almost optimal adaptation in a larger frame
of reference, e.g., if we take into account the metabolic costs for increased perceptual
accuracy. We will not engage in such speculation here. Rather we will explore the consequences
of perceptual limitations on processes of meaning evolution in a suitable formal framework. In
other words, in order to address Lipman's challenge seriously, and not just hand-wave it away
by appeal to the naturalness of vagueness, it needs to be spelled out how exactly confusability
of stimuli could come into play in a process of meaning evolution and how it could lead to
vague but by-and-large informative signal meaning. The formal model that this paper introduces
does exactly that. But it also does more. We find that confusability of stimuli can regularise
and systematise evolving meaning. This suggests a possibly advantageous side-effect that a
natural cause of vagueness may have and that might compensate some of the disadvantages that
vagueness may have for communicative efficiency \citep[c.f.][]{OConnor2013:The-Evolution-o}.

@@

The next section introduces the background against which the work presented here can be
appreciated. Section~\ref{sec:repl-diff-dynam} introduces a generalisation of the replicator
dynamic which is derived from the idea that agents imitate other agents' behaviour while
possibly confusing similar states. This imprecise imitation dynamic is explored in
Section~\ref{sec:exploring-rdd}. Section~\ref{sec:discussion} reflects and compares our
approach to others. Appendix~\ref{sec:formal-stuff} provides formal detail.

@@

Signalling games, as introduced by \citet{Lewis_1969:Convention}, have
a sender and a receiver. The sender knows the true state of the world,
but the receiver does not. The sender can select a signal, or message,
to reveal to the receiver, who then chooses an act. In Lewis' games,
if the receiver chooses the act that corresponds to the actual state,
the play is a success, otherwise a failure. Certain regular
combinations of sender signalling and receiver reaction make messages
meaningful, in the sense that their use is correlated systematically
to certain states or acts. To investigate the conditions under which
such meaning-generating behaviour can evolve is a highly interesting
topic that we are only beginning to fully understand
\citep[e.g.][]{Warneryd1993:Cheap-Talk-Coor,BlumeKim1993:Evolutionary-St,Huttegger2007:Evolution-and-t,Pawlowitsch2008:Why-Evolution-d,Barrett2009:The-Evolution-o,HutteggerSkyrms2010:Evolutionary-Dy,Skyrms2010:Signals}.

Similarity-maximising (short: sim-max) games are a variation of Lewis' games where the
receiver's actions are equated with the state space (one can think of the actions as choosing
states) and different states are allowed to be more or less similar to one another. While
Lewis' games treated communicative success as a matter of black and white, sim-max games allow
for shades of grey: the more similar the receiver's interpretation is to the actual state, the
better. Signalling games with utility-relevant similarities in the state space are fairly
standard in economics
\citep[e.g.][]{Spence1973:Job-market-sign,CrawfordSobel1982:Strategic-Infor}, but have received
particular attention in a more philosophical context for reasons that will become clear
presently.

Formally, a sim-max game consists of a set of states $\States$, a set of messages $\Messgs$
typically with much fewer messages than states, a probability distribution
$\Pr \in \Delta(\States)$ such that $\Pr(\state)$ gives the probability that state $\state$
occurs, a similarity metric\footnote{A metric is a function of distance between any two points
  in a given space. It should satisfy certain axioms that ensure behaviour that one would
  intuitively expect from the words ``similarity'' and ``distance'' alone, but these details do
  not matter for the purposes of this paper. The assumption that state similarity forms a
  metric is conceptually loaded, but we follow the literature here and conceive of it as a
  first and pragmatic simplification, possibly to be dispensed with later.}
  % : similarity should always be non-negative, zero only for the distance
  % between an element and itself, symmetric, and satisfy the triangle inequality that
  % $\similarity(x,z) \le \similarity(x,y) + \similarity(y,z)$
on states $\similarity \mycolon \States \times \States \rightarrow \mathds{R}$ such that
$\similarity(\mystate{1}, \mystate{2})$ is the (physical) similarity between $\mystate{1}$ and
$\mystate{2}$, and a utility function
$\Utils \mycolon \States \times \States \rightarrow \mathds{R}$ such that
$\Utils(\mystate{1}, \mystate{2})$ is the payoff for sender and receiver for a play with actual
state $\mystate{1}$ and receiver interpretation $\mystate{2}$. We identify the receiver's acts
with the states of the world, so that the game is one of guessing the actual state, so to
speak. For the modelling purposes of this paper, we make the simplifying assumption that
$\States$ is a set of points in Euclidean space, whose closeness to each other tracks physical
similarity. Perceived similarity, where it is necessary, would be a monotonic function of
physical similarity. Likewise, the utility function should be a monotonically increasing
function of physical similarity.\footnote{Section~\ref{sec:setting-stage} motivates particular
  choices of similarity and utility functions that we will explore in more detail.}
Non-probabilistic sender behaviour can be represented by a pure strategy
$\Spure \in \Messgs^\States$, which deterministically defines which message would be used for
each state. Similarly, a pure receiver strategy is a function $\Rpure \in \States^\Messgs$.

@@

It is intuitive to think that linguistic and conceptual categories are orderly in such a
manner.  For example, if you consider two people `tall', one with a height of 2m and another with a
height of 2.2m, it would be difficult to defend not considering a person with a height
of 2.1m `tall' as well. Another example where this intuition is additionally supported by empirical data is that of colour categorisation. The World Color Survey
project~\citep{CookKay2005:The-World-Color,WCS} collected colour naming data for 110 unwritten
languages of 45 language families.  For the great majority of these languages a pattern can be
observed: basic colour terms are by and large convex~\citep{Regier07,Jager2010}. It
is premature to argue that these observations can be extended to \emph{all} cases of
categorisation. However, for the cases where it does apply, the result of
\citet{JagerMetzger2011:Voronoi-Languag} is interesting because it demonstrates that signalling
can impose this kind of orderly categories on a metric space without that being the ulterior
purpose of it all.%
\footnote{For the concrete case of colour categorisation, see also
  \citet{JagerRooijvan-Rooij2007:Language-Struct} and \citet{Correia2015}}
Finally, these considerations are in line with a prominent
school of thought in comparative linguistics which also assumes that more abstract conceptual domains
(e.g., spatial-topological relations, temporal reference, or the meanings expressible by
indefinite pronouns) are preferably carved up by the languages of the world in such a way
that meanings are connected regions on a so-called \emph{semantic map}
\citep[e.g.][]{Croft2003:Typology-and-Un,Haspelmath2003:The-geometry-of,LevinsonMeira2003:Natural-concept}.

There are at least two potential ways of interpreting the signalling setup.
Sender and receiver can be distinct
entities, whose purpose is to communicate effectively about the actual
state. In that case, evolving Voronoi languages would explain why
\emph{linguistic categories} are well-behaved and orderly in the way
they appear to be. More abstractly, sender and receiver can also be
thought of as distinct modules in a single system, where the first
module must discretise the information it is fed by selecting a small
sample of, suggestively, category labels. These are passed to a second
module that tries to decode the original information. In this case,
evolving Voronoi languages would explain why \emph{conceptual
  categories} are well-behaved and orderly in the way that they appear
to be.
Seen in this light, sim-max games may provide a foundation to approaches in cognitive semantics
that rely on the notion of conceptual spaces.
\citet[][70--77]{Gardenfors2000:Conceptual-Spac}, for example, has prominently argued that
natural categories are convex regions in conceptual space. If the conceptual space has a
suitable metric, convex categories can be derived from a set of prototypes. The category
corresponding to prototype $p$ is the set of points that are more similar to $p$ than to any
other. In this way, \citeauthor{Gardenfors2000:Conceptual-Spac} argues, an efficient categorisation
system can be obtained: storing the prototypes lets us recover the categories without having to
store each category's extension. However, what is left unexplained is where the prototypes
come from, and why we would not see just any distribution of prototypes as an equally efficient
classification system. This is where sim-max games can contribute a principled approach to
deriving, in an independent way, not only convex categories but also prototypical exemplars
belonging to them.  These ideas, and more, are developed further by
\citet{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag,OConnor2014-OCOEPC},
among others.

@@

This outline of an approach to categorisation using sim-max games
leaves many problems unaddressed. One of them is that natural
categories for continuously variable stimuli usually do not have
unique, point-valued prototypes and clear category boundaries. We
would like to account for the possibility of such vagueness, in
particular: (i) clear positive examples of a vague category should
show a gradient transition to clear negative examples;
(ii) prototypes should likewise be gradient regions, peaking at the
centre of the vague category they represent.

@@

Alternative approaches are taken by, for example, \citet{FrankeJager2010:Vagueness-Signa} and
\citet{OConnor2013:The-Evolution-o}, who show different ways how the above desiderata can be
met by evolving strategies in sim-max games. To illustrate what vague signalling would look
like, let us briefly consider a sim-max game with six equiprobable states and two messages and
what its equilibria would be like \citep[see][for a more thorough discussion of equilibria of
sim-max games]{Jager2007:The-Evolution-o,OConnor2013:The-Evolution-o}. We assume that utilities
are linearly decreasing with decreasing similarity. Figure~\ref{fig:EquilibriaVagueness} shows
three pairs of sender and receiver strategies. States are arranged according to their
similarity: the closer they are to each other, the more similar they are. The pair in
Figure~\ref{fig:ineffNonEq} is not an equilibrium, because the sender's non-convex use of
signals is sub-optimal given the receiver's behaviour. Namely, if \mymessg{1} is interpreted as
\mystate{2} and \mymessg{2} as \mystate{4}, then the sender would get a higher payoff from
sending \mymessg{1} in \mystate{3} than from sending \mymessg{2}, because (by assumption)
\mystate{3} is more similar to \mystate{2} than \mystate{5} is. In contrast,
Figure~\ref{fig:efficientEq} shows a maximally efficient equilibrium. This is a partial pooling
equilibrium in the sense that the speaker uses the same message for several states. Partial
pooling equilibria can be less inefficient than other non-pooling equilibria, if there are
enough messages. The strategy pair of Figure~\ref{fig:efficientEq} is maximally efficient for
the two-message case. So, while partial pooling may entail inefficiency in some sense, and
while partial pooling can hamper the evolution of maximally efficient signal use
\citep[e.g.][]{Hutteger:2007_Evol_Indicatives_Imperatives,Pawlowitsch2008:Why-Evolution-d,HutteggerSkyrms2010:Evolutionary-Dy},
this is orthogonal to our concerns about vagueness. A regular and natural yet vague signal use
would look like the pair in Figure~\ref{fig:vagueUse}. This is not an equilibrium, but it gets
close, so to speak. It shows smooth transitions across similar states at the boundaries of
categories and across acts around the most prototypical instances (as indicated by decreasing
thickness of arrows in Figure~\ref{fig:vagueUse}).

@@

Our conceptual starting point is the widely shared conviction that language is shaped by at
least two forces, which may, on occasion, pull in opposite direction. On the one hand, there is
functional pressure towards efficient communication. On the other hand, there is systematic
error, noise or imprecision in the transmission of linguistic behaviour, knowledge or traits. As
an example for the latter, consider a child learning syntactic rules from a parent
generation. The child must infer these unobservable rules from observable speech. Inductive
biases may influence which syntactic rules are likely to be inferred from (finite) parental
input. Over the course of many generations, the effects of such biases can, in a manner of
speaking, ``accumulate'' and lead to surprising results, such as the evolution of compositional
form-meaning mappings or the use of regular recursive syntactic structure. This can happen when
the effects of transmission biases are isolated, as in iterated learning models
\citep{KirbyHurford2002:The-Emergence-o,SmithKirby2003:Iterated-Learni,GriffithsKalish2007:Language-Evolut,KirbyGriffith2014:Iterated-Learni},
or when they interact with functional pressure towards efficient communication, e.g., as
formalised in the replicator mutator dynamic
\citep[e.g.][]{NowakPlotkin2000:The-Evolution-o,NowakKomarova2001:Evolution-of-Un}.

The emphasis of previous models that studied the effects of transmission infidelity on the
evolution of language has been on inductive biases and the systematicity, compression and
regularisation that they can introduce. Here, we would like to show that transmission noise of
a different kind can lead to regularisation as well and also give rise to vague meaning. We
find that shared perceptual biases that perturb the transmission of successful signalling
behaviour can regularise, facilitate and accelerate the evolution of meaning conventions, albeit
at the cost of vagueness. Concretely, we formalise the expected change in the behaviour of a
population of agents that try to imitate other agents' signalling behaviour. We assume, however,
that both realisation of behaviour, as well as observability of others' behaviour, is
systematically perturbed by noise. The resulting population-level dynamic generalises the
replicator dynamic \citep{TaylorJonker1978:Evolutionary-St} in its interpretation as a cultural
evolutionary dynamic based on imitation
\citep[e.g.][]{Helbing1996:A-Stochastic-Be,Schlag1998:Why-Imitate-and}. The inclusion of
confusability of stimuli does not undermine the possibility of evolving communicative signalling
behaviour. Instead, it leads to the evolution of vague meanings. It moreover accelerates the
emergence of communicative signalling because it unifies and regularises evolutionary outcomes,
making it appear as if agents were applying inductive biases or generalising over partial
observations, when in truth this is the sole effect of confusion of similar stimuli.

@@

Signalling agents can adapt their dispositions to act, given some feedback about their past
success, in multiple ways. Usually, we would assume that changes in behaviour should, at least
on average, lean towards increasing chances of communicative success. Such behavioural
adaptations can be described at different levels of abstraction. At the level of individual
agents, we can picture a more or less idealised process of how each agent adapts dispositions
for future actions based on various pieces of information available to the agent. More
abstractly, at the level of a population of agents, we can describe how average behavioural
dispositions will evolve. The population-level perspective abstracts over small stochastic
fluctuations and zooms in on the general tendency or direction of evolution that ensues from
behaviour at the agent-level.

To better understand the interaction of confusability of stimuli and selective pressure towards
successful communication, we look at a population-level dynamic that describes the most likely
evolutionary path of a population of signalling agents who imitate other agents' behaviour but
are liable to confuse states for one another. In the special limiting case where state
confusability vanishes, the process is just the well-known replicator dynamic
\citep{TaylorJonker1978:Evolutionary-St}, which should therefore be briefly reviewed first.

\subsection{Replicator dynamic in behavioural strategies}
\label{sec:repl-dynam-behav}

Fix a sim-max game with finite states $\States$ and messages $\Messgs$. As usual, we assume
that the receiver chooses states in $\States$ in response to messages. Let
$\Pr(\cdot) \in \Delta(\States)$ be the prior distribution over states and
$\Util \mycolon \States \times \States \rightarrow \mathds{R}$ the utility function shared by
senders and receivers in the population. A \emph{behavioural strategy} is a function that maps
an agent's choice points to a probability distribution over available
choices.\footnote{\label{fn:mixedStrats}Our focus is on behavioural strategies, not \emph{mixed
    strategies}, i.e., probability distributions over functions from each choice point to an
  act, such as $s \in \Delta(\Messgs^{\States})$.  Dynamics on behavioural strategies assume
  that agents can adapt their behaviour locally, i.e., independently at each choice point. Our
  focus on behavioural strategies greatly reduces the complexity of the dynamic and simplifies
  numerical simulations. But it also seems the more plausible choice for imitation-based update
  protocols of the kind we consider here: agents only observe how, on some occasion, some other
  agent behaved in \emph{one} particular situation, not how that agent would behave in
  \emph{all} relevant choice situations; they imitate the use of a single word, so to speak,
  not a whole lexicon. (See \citet{Cressman2003:Evolutionary-Dy} for more on the difference
  between dynamics on mixed or behavioural strategies.) It is in this respect that the cultural
  evolutionary dynamic introduced here differs most visibly from the replicator mutator dynamic
  \citep[e.g.][]{NowakPlotkin2000:The-Evolution-o,NowakKomarova2001:Evolution-of-Un}, which
  operates on mixed strategies and is motivated by assumptions of (asexual) biological
  inheritance with transmission infidelity.} The sender's behavioural strategies are
functions $\Sstrat \in \Delta(\Messgs)^\States$, thus mapping each state $\state \in \States$
to a probability of each message $\messg \in \Messgs$ being sent in $\state$; the receiver's
are functions $\Rstrat \in \Delta(\States)^\Messgs$, thus mapping each message
$\messg \in \Messgs$ to a probability of each interpretation $\state \in \States$ being chosen
in response to $\messg$. Although behavioural strategies are probabilistic, evolutionary
modelling usually imagines that every individual agent has a non-probabilistic
strategy. Behavioural strategies then capture average population behaviour. Assuming a virtually
infinite population, the number $\Sstrat(\messg \probbar \state)$, for instance, is then the
probability that a randomly sampled sender would send message $\messg$ if the actual state was
$\state$. Similarly, $\Rstrat(\state \probbar \messg)$ is then the probability with which a
randomly sampled receiver interprets $\messg$ as $\state$. A play of a single evolutionary game
with behavioural strategies is illustrated in Figure~\ref{fig:sim-max-illustration}.
